A systematic search for transiting planets in the \KT\ data

### 2015-02-12 ms.tex
a bit of reordering in discussion
editing pipeline intro and linear search

\section{Joint transit \& variability model}
\label{sec:model}

@@

To avoid over-fitting, in our pipeline we simultaneously fit for the transit
signal and the trends using a rigid model for the signal and a relatively
flexible model for the noise.
Specifically, we model the light curve as being generated by linear
combination of 150 ELCs and a ``box'' transit model at a specific period,
phase, and duration.
The mathematical details are given in \app{math} but in summary, since the
model is linear, we can analytically compute the likelihood
function---conditioned on a specific period, phase, and duration---for the
depth \emph{marginalized over the systematics model}.
This computation is expensive but, as described in the following sections, it
is possible to scale the method to a \KT-scale dataset.

@@

In theory, search could proceed by evaluating the model described above on a
fine three-dimensional grid in period, phase, and duration.
In practice, this is computationally intractable for any grids of the
required size.
Instead, we can compute the values on this grid approximately, but at very
high precision, using a two-step procedure that is much more computationally
efficient.

The specific quantity that we will be searching is the likelihood for the
light curve of star $n$ given a set of transit parameters:
\begin{eqnarray}
p(\{\flux\}_n\,|\,\period,\,\phase,\,\duration,\,\depth) \quad.
\end{eqnarray}
We will make the simplifying assumption that each transit is independent
because it will be approximately satisfied for all but the shortest periods
and it leads to a huge computational advantage.
Under this assumption, this likelihood function can be rewritten as
\begin{eqnarray}\eqlabel{indtran}
p(\{\flux\}_n\,|\,\period,\,\phase,\,\duration,\,\depth) &\approx&
\prod_{m=1}^{M(\period,\,\phase)}
    p(\{\flux\}_n\,|\,\transittime_m(\period,\,\phase),\,\duration,\,
                    \depth)
\end{eqnarray}
where $\transittime_m(\period,\,\phase)$ is the time of the $m$-th
transit given the period $\period$ and reference time $\phase$.
\Eq{indtran} can be efficiently computed for many periods and phases if we
first compute a set of likelihood functions for single transits on a fine grid
in $\transittime_l$
\begin{eqnarray}\eqlabel{singletransit}
\left \{ p(\{\flux\}_n\,|\,\transittime_l,\,\duration_k,\,\depth)
\right\}_{l=1,\,k=1}^{L,\,K}
\end{eqnarray}
then use these results as a lookup table.

In the remainder of this section, we give more details about each step of the
search procedure but in summary, it breaks into three main steps: linear
search, periodic search, and vetting.
In the {\bf linear search} step, we evaluate the likelihood function in
\eq{singletransit} on a two-dimensional grid, coarse in transit duration
$\duration_k$ and fine in transit time $\transittime_m$.
Then in the {\bf periodic search} step, we use this two-dimensional grid to
approximately evaluate the likelihood (\eqalt{indtran}) for a
three-dimensional grid of periodic signals.
Then, we run a peak detection algorithm on this grid that discards peaks that
share transits with a shorter period candidate and signals with substantially
varying transit depths.
These transit candidates are then passed along for machine and human {\bf
vetting}.

@@

The linear search involves hypothesizing a transit signal on a
two-dimensional grid in transit time and duration.
For each point in the grid, we use the model described in the previous
section to evaluate \emph{the likelihood function} of the transit depth.
Because the model is linear, the likelihood function for the depth
(marginalized over the model of the systematics) is a Gaussian with
analytic amplitude $L$, mean $\bar{\depth}$, and variance
$\delta\bar{\depth}^2$ derived in \app{math}.
Therefore in the linear search, we save these three numbers for a
two-dimensional grid in transit time \transittime\ and duration \duration.
The transit time grid spans the duration of Campaign 1 with half hour spacing
and we choose to only test three durations: 1.2, 2.4, and 4.8 hours.
\Fig{linear} shows the mean transit depth $\bar{\depth}$ as a function of
transit time \transittime\ for a light curve with an injected transit signal
with transits at the times indicated by the vertical green bars.
The computational complexity of this step is $\mathcal{O}(M\,K^3)$ where $M$
is the number of points in the grid and $K$ is the number of data points.

@@

\clearpage
\appendix

>>>>

\section{Joint transit \& variability model}
\sectlabel{model}

@@

To avoid over-fitting, in our pipeline we simultaneously fit for the transit
signal and the trends using a rigid model for the signal and a relatively
flexible model for the noise.
Specifically, we model the light curve as being generated by linear
combination of 150 ELCs and a ``box'' transit model at a specific period,
phase, and duration.
The mathematical details are given in \app{math} but in summary, since the
model is linear, we can analytically compute the likelihood
function---conditioned on a specific period, phase, and duration---for the
depth \emph{marginalized over the systematics model}.
The signal-to-noise of this depth measurement can then be used as a quality
of fit metric or candidate selection scalar.
This computation is expensive but, as described in the following sections, it
is possible to scale the method to a \KT-scale dataset.

@@

In principle, the search for transit signals simply requires evaluating the
model described above on a fine three-dimensional grid in period, phase, and
duration, and then detecting high significance peaks in that space.
In practice, this is computationally intractable for any grids of the
required size.
Instead, we can compute the values on this grid approximately, but at very
high precision, using a two-step procedure that is much more computationally
efficient.

Specifically, we must evaluate the likelihood function for the light curve of
star $n$ given a period \period, reference transit time \phase, and duration
\duration
\begin{eqnarray}\eqlabel{full-likelihood}
p(\{\flux\}_n\,|\,\period,\,\phase,\,\duration,\,\depth) \quad.
\end{eqnarray}
We make the simplifying assumption that each transit enters this quantity
independently because it is approximately satisfied for all but the shortest
periods and it leads to a huge computational advantage.
Under this assumption, this likelihood function can be rewritten as
\begin{eqnarray}\eqlabel{indtran}
p(\{\flux\}_n\,|\,\period,\,\phase,\,\duration,\,\depth) &=&
\prod_{m=1}^{M(\period,\,\phase)}
    p(\{\flux\}_n\,|\,\transittime_m(\period,\,\phase),\,\duration,\,
                    \depth)
\end{eqnarray}
where $\transittime_m(\period,\,\phase)$ is the time of the $m$-th
transit given the period $\period$ and reference time $\phase$.
\Eq{indtran} can be efficiently computed for many periods and phases if we
first compute a set of likelihood functions for single transits on a grid
in $\transittime_l$ and duration $\duration_k$
\begin{eqnarray}\eqlabel{singletransit}
\left \{ p(\{\flux\}_n\,|\,\transittime_l,\,\duration_k,\,\depth)
\right\}_{l=1,\,k=1}^{L,\,K} \quad.
\end{eqnarray}
Then, we can use these results as a look-up table to approximately evaluate
\eq{full-likelihood}.

In the remainder of this section, we give more details about each step of the
search procedure but in summary, it breaks into three main steps: linear
search, periodic search, and vetting.
In the {\bf linear search} step, we evaluate the likelihood function in
\eq{singletransit} on a two-dimensional grid, coarse in transit duration
$\duration_k$ and fine in transit time $\transittime_m$.
Then in the {\bf periodic search} step, we use this two-dimensional grid to
approximately evaluate the likelihood (\eqalt{indtran}) for a
three-dimensional grid of periodic signals.
Then, we run a peak detection algorithm on this grid that discards signals
with substantially varying transit depths.
These transit candidates are then passed along for machine and human {\bf
vetting}.

@@

The linear search requires hypothesizing a set of transit signals on a
two-dimensional grid in transit time and duration.
For each point in the grid, we use the model described in \sect{model} to
evaluate \emph{the likelihood function} for the transit depth at that time
and duration.
Since the model is linear, the likelihood function for the depth (marginalized
over the model of the systematics) is a Gaussian with analytic amplitude $L$,
mean $\bar{\depth}$, and variance $\delta\bar{\depth}^2$ derived in
\app{math}.
Therefore in the linear search, we save these three numbers for a
two-dimensional grid in transit time $\transittime_l$ and duration
$\duration_k$.
The transit time grid spans the duration of Campaign 1 with half hour spacing
and we choose to only test three durations: 1.2, 2.4, and 4.8 hours.
\Fig{linear} shows the mean transit depth $\bar{\depth}$ as a function of
transit time \transittime\ for a light curve with an injected transit signal
with transits at the times indicated by the vertical green bars.

% The computational complexity of this step is $\mathcal{O}(M\,K^3)$ where $M$
% is the number of points in the grid and $K$ is the number of data points.

@@

\appendix
